---
title: "BT2103_ProblemSet3_A0236437B"
author: "Lo Zhi Hao"
date: "2022-09-04"
output: html_document
---

```{r setup}
library(dplyr)
library(readr)
library(linprog)
```

# BT2103 Tutorial # 3, September 7, 2022

The dataset for this tutorial is a subset of the Credit Approval Data from UCI Machine
Learning Repository:

[link](https://archive.ics.uci.edu/ml/datasets/credit+approval)

Only two attributes (3 and 15) and a subset of the samples will be used in this exercise. The
two continuous attributes are:

• x1: debt (original attribute 3)

• x2: income (original atrribute 15)

The two classes in the data (attribute 16) have been labeled as:

• Class 0: credit denied (original value ”-”)

• Class 1: credit approved (original value ”+”)

1. Download the dataset ”crx-subset.txt” and the file ”LPinMPS.txt” from Luminus Tutorial folder.

```{r 1}
## test

crx_data <- read.delim("crx-subset.txt", header = FALSE, dec = ".")
summary(crx_data)
head(crx_data)

LP_data <- read.delim("LPinMPS.txt", header = FALSE, sep = "\t", dec = ".")
summary(LP_data)
head(LP_data)
```

2. We are finding a linear decision boundary to separate the Class 0 samples from the
Class 1 samples by solving the LP:

max α−β

subject to:

w1x1 +w2x2 ≥ α for all approved samples Class 1

w1x1 +w2x2 ≤ β for all denied samples Class 0

w2 = 1

The linear programming problem with the relavant data in the file crx-subset.txt has
been formatted as an MPS file. MPS file format is useful when the linear program
is sparse: it has many zero coefficients in the cost vector, or rhs vector, or in the
constraint matrix. More information about the MPS file can be found in

• [link](https://rdrr.io/cran/linprog/man/readMps.html)

• [link](https://cran.r-project.org/web/packages/linprog/linprog.pdf)

• [link](http://web.mit.edu/lpsolve_v5520/doc/mps-format.html)

Solve the linear program presented in LPinMPS.txt by installing and loading the
linprog package and reading the MPS input data file LPinMPS.txt:

lp <- readMps("LPinMPS.txt",solve=TRUE,maximum=TRUE)

1
What are the optimal values for w1,w2,α and β?

```{r q2}
## solving linear programming problem 
lp <- readMps("LPinMPS.txt",solve=TRUE,maximum=TRUE)
summary(lp)
head(lp)

## Optimum values for w1 is 8.97135, w2 = 1, BETA = 594.63293, ALPHA = 308.97135
```

3. What is the purpose of having the constraint

w2 = 1

in the LP formulation? You may answer this question by

• removing the constraints w2 ≥ 1 and −w2 ≥ −1 from the file LPinMPS.txt,

• saving the file to LPinMPS1.txt,

• re-solving the LP by calling readMps again.

```{r q3}
## using from file LPinMPS1_q3.txt

lp <- readMps("LPinMPS_q3.txt", solve = TRUE, maximum = TRUE)
summary(lp)
head(lp)

## w1, w2, ALPHA, BETA are all 0 
```

4. Plot the data points in crx-subset.txt and the following decision boundaries:

w1x1 +w2x2 = θ

for three values of θ = β,(α+β)/2 and α.

```{r q4}
## using table to store data "crx_table"
crx_table <- read.table("crx-subset.txt", header = FALSE)
crx_v1 <- crx_table$V1
crx_v2 <- crx_table$V2
crx_v3 <- crx_table$V3
head(select(crx_table, -"V3"))
crx_table_w1w2 <- select(crx_table, -"V3")
plot(crx_table_w1w2, ylim = c(0, 1000), xlim = c(0, 80), xlab = "Weight 1", ylab = "Weight 2", col = ifelse(crx_table$V3 == 0, "red", "blue"), pch = ifelse(crx_table$V3, 15, 16))

## θ = β,(α+β)/2 and α
intercept1_beta <- 594.63293
intercept2_alphabeta <- (308.97135 + 594.63293) / 2
intercept3_alpha <- 308.97135	
slope_line <- -8.97135	
abline(intercept1_beta, slope_line, col = "green")
abline(intercept2_alphabeta, slope_line, lty = 3)
abline(intercept3_alpha, slope_line, col = "orange")
```

5. The decision rule is as follows:

If w1x1 +w2x2 < θ, predict Class 0 (negative),
else predict Class 1 (positive)
Compute the false positive rate (FPR) and true positive rate (TPR) for each of the 3
threshold values θ = β+10−4
,(α+β)/2 and α−10−4
.

```{r q5}
## use a for loop to loop through the values in crx_table then produce results accordingly 

alpha	<- 308.971348	
beta <-	594.632926
w1 <-	8.971348
w2 <- 1

x <- crx_v1 
y <- crx_v2 
z <- crx_v3
n <- length(x)

delta = 1e-04 # is added/subtracted to avoid samples falling exactly on the hyperplanes
vtresh <- c(alpha-delta, (alpha + beta)/2, beta+delta)
ypred <- w1*x + w2*y
TPRATE <- 0
FPRATE <- 0
j <- 1

 while (j <= 3) {
i <- 1
tn <- 0
fn <- 0
tp <- 0
fp <- 0
tpr <- 0
fpr <- 0
while (i <= n) {
thresh <- vtresh[j]
if (ypred[i] < thresh && z[i] == 0) { tn <- tn + 1}
else if (ypred[i] < thresh && z[i] == 1) { fn <- fn + 1}
else if (ypred[i] >= thresh && z[i] == 1) { tp <- tp +1}
else {fp <- fp + 1}
i <- i + 1}
cat("Threshold = " ,thresh, " TN FN TP FP ",tn,fn,tp,fp,"\n")
j1 <- 5 - j
tpr <- tp/(tp + fn)
fpr <- fp/(tn + fp)
FPRATE[j1] <- fpr
TPRATE[j1] <- tpr
cat("Threshold = " ,thresh, " TPR FPR ",tpr,fpr,"\n\n")
j <- j + 1
 }

```

6. Plot the ROC curve and compute the area under this curve.

```{r q6}
TPRATE[1] = 0
FPRATE[1] = 0
TPRATE[5] = 1
FPRATE[5] = 1
area = 0

i <- 2
while (i <= 5)
{
b1 <- FPRATE[i]
b2 <- FPRATE[i-1]
h1 <- TPRATE[i]
h2 <- TPRATE[i-1]
area <- area + 0.5*(b1-b2)*(h1 + h2)
i <- i + 1
}
cat("Area under ROC curve = " ,area,"\n")
```

7. What will be the next step of the MSM (Multisurface Method)?


```{r q7}
## Next step is to solve a subset of the original thing, 
## which is the part between alpha and beta lines. 
```
