---
title: "BT2103_PS2_A0236437B_LoZhiHao"
author: "Lo Zhi Hao"
date: "2022-08-29"
output: html_document
---
## Setup 

```{r setup, results = FALSE}
## install.packages("lpSolve")
## install.packages("lpSolveAPI")
## install.packages("linprog")
library("lpSolve")
library("lpSolveAPI")
library("linprog")
```

## Question 1 

(a) Explicitly show all 10 constraints in the linear program.

```{r q1a}
## min a
## Subject to:
## (For default samples: )
## w1*10 + w2*24 ≥ c − a 
## w1*25 + w2*35 ≥ c − a 
## w1*3 + w2*40 ≥ c − a 
## w1*10 + w2*30 ≥ c − a 
## w1*30 + w2*20 ≥ c − a 
## w1*18 + w2*20 ≥ c − a 
## (For non-default samples: )
## w1*2 + w2*28 ≤ c + a
## w1*5 + w2*21 ≤ c + a
## w1*4 + w2*30 ≤ c + a
## w1*0 + w2*33 ≤ c + a
## a ≥ 0
```

(b) Use R (or any other LP solver) to find the optimal solution. Fix the value of c
in the LP formulation c = 100.

```{r test}
## test 
cvec <- c(8,12)
bvec <- c(200,8000,-100,-200)
names(cvec) <- c("X1", "X2")
names(bvec) <- c(" Labor ", "Budget", "X1lower", "X2lower")
Amat <- rbind( c(0.3,0.5), c(10,15), c(-1,0), c(0,-1))
res <- solveLP(cvec, bvec, Amat, const.dir = c("<=", "<=", "<=", "<="), maximum = TRUE)
print(res)
```

```{r q1b}
## min 0 * X1 + 0 * X2 + 1 * a
## Subject to:
## (For default samples: )
## w1*10 + w2*24 + a ≥ 100 
## w1*25 + w2*35 + a ≥ 100 
## w1*3 + w2*40 + a ≥ 100
## w1*10 + w2*30 + a ≥ 100  
## w1*30 + w2*20 + a ≥ 100  
## w1*18 + w2*20 + a ≥ 100 
## (For non-default samples: )
## w1*2 + w2*28 - a ≤ 100 
## w1*5 + w2*21 - a ≤ 100
## w1*4 + w2*30 - a ≤ 100 
## w1*0 + w2*33 - a ≤ 100 
## a ≥ 0

## q1b

cvec <- c(0, 0, 1)
bvec <- c(100, 
          100, 
          100, 
          100, 
          100, 
          100, 
          100, 
          100, 
          100, 
          100)
names(cvec) <- c("X1", "X2", "a")
names(bvec) <- c("Default1", 
                 "Default2", 
                 "Default3", 
                 "Default4", 
                 "Default5", 
                 "Default6", 
                 "Non-Default1", 
                 "Non-Default2", 
                 "Non-Default3", 
                 "Non-Default4")
Amat <- rbind(c(10, 24, 1), 
              c(25, 35, 1), 
              c(3, 40, 1),
              c(10, 30, 1),
              c(30, 20, 1),
              c(18, 20, 1),
              c(2, 28, -1),
              c(5, 21, -1),
              c(4, 30, -1),
              c(0, 33, -1))
operator <- c(">=",
              ">=",
              ">=",
              ">=",
              ">=",
              ">=",
              "<=",
              "<=",
              "<=",
              "<=")
res <- solveLP(cvec, bvec, Amat, const.dir = operator, maximum = FALSE)
print(res)


```

```{r q1b way2}
f.obj <- c(0, 0, 1)

f.con <- matrix(c(10, 24, 1,
                  25, 35, 1,
                  3, 40, 1,
                  10, 30, 1,
                  30, 20, 1,
                  18, 20, 1,
                  2, 28, -1,
                  5, 21, -1,
                  4, 30, -1,
                  0, 33, -1), nrow = 10, byrow = TRUE)
f.dir <- c(">=",
           ">=",
           ">=",
           ">=",
           ">=",
           ">=",
           "<=",
           "<=",
           "<=",
           "<=")

f.rhs <- c(rep(100,10))

lp.solution <-  lp("min", f.obj, f.con, f.dir, f.rhs, compute.sens = TRUE)

lp.solution$solution
```

(c) Are the samples linearly separable? Plot the data to confirm your answer

```{r q1c}
## q1c
## Creating data frame

## Default
CreditCardBalance <- c(10, 25, 3, 10, 30, 18)
Age <- c(24, 35, 40, 30, 20, 20)
cust <- c("Cust1","Cust2","Cust3","Cust4","Cust5","Cust6")

data <- data.frame(CreditCardBalance, Age)
plot(data, col = "red", pch = 1, xlim = c(0, 50), ylim = c(0, 50))
text(CreditCardBalance, Age, cust, pos = 1)

## Non default
CreditCardBalance_nonDefault <- c(2, 5, 4, 0)
Age_nonDefault <- c(28, 21, 30, 33)
cust_nonDefault <- c(cust <- c("Cust7","Cust8","Cust9","Cust10"))
points(x = CreditCardBalance_nonDefault, y = Age_nonDefault, col = "blue", pch = 2)
text(CreditCardBalance_nonDefault, Age_nonDefault, cust_nonDefault, pos = 1)

## Yes I think the samples are linearly separable

intercept <- 100 / 1.45985
slope <- -13.86861 / 1.45985
abline(intercept, slope)

## However, there will be at least one error, as three points lie on the curve, but they are classified as different categories. 
```

## Question 2

(a) Find the best hyperplane that minimizes the maximum deviation.

```{r q2a}
## min 0 * X1 + 0 * X2 + 1 * a
## Subject to:
## (For default samples: )
## w1*10 + w2*24 + a ≥ 100 
## w1*25 + w2*35 + a ≥ 100 
## w1*3 + w2*40 + a ≥ 100
## w1*10 + w2*30 + a ≥ 100  
## w1*30 + w2*20 + a ≥ 100  
## w1*18 + w2*20 + a ≥ 100 
## (For non-default samples: )
## w1*2 + w2*28 - a ≤ 100 
## w1*5 + w2*21 - a ≤ 100
## w1*4 + w2*30 - a ≤ 100 
## w1*10 + w2*33 - a ≤ 100 
## a ≥ 0

## q2a

cvec <- c(0, 0, 1)
bvec <- c(100, 
          100, 
          100, 
          100, 
          100, 
          100, 
          100, 
          100, 
          100, 
          100)
names(cvec) <- c("X1", "X2", "a")
names(bvec) <- c("Default1", 
                 "Default2", 
                 "Default3", 
                 "Default4", 
                 "Default5", 
                 "Default6", 
                 "Non-Default1", 
                 "Non-Default2", 
                 "Non-Default3", 
                 "Non-Default4")
Amat <- rbind(c(10, 24, 1), 
              c(25, 35, 1), 
              c(3, 40, 1),
              c(10, 30, 1),
              c(30, 20, 1),
              c(18, 20, 1),
              c(2, 28, -1),
              c(5, 21, -1),
              c(4, 30, -1),
              c(10, 33, -1))
operator <- c(">=",
              ">=",
              ">=",
              ">=",
              ">=",
              ">=",
              "<=",
              "<=",
              "<=",
              "<=")
res <- solveLP(cvec, bvec, Amat, const.dir = operator, maximum = FALSE)
print(res)

```

(b) How many data samples have non-zero deviations? What is the maximum deviation?

```{r q2b}

CreditCardBalance <- c(10, 25, 3, 10, 30, 18)
Age <- c(24, 35, 40, 30, 20, 20)
cust <- c("Cust1","Cust2","Cust3","Cust4","Cust5","Cust6")

data <- data.frame(CreditCardBalance, Age)
plot(data, col = "red", pch = 1, xlim = c(0, 50), ylim = c(0, 50))
text(CreditCardBalance, Age, cust, pos = 1)

## Non default
CreditCardBalance_nonDefault <- c(2, 5, 4, 10)
Age_nonDefault <- c(28, 21, 30, 33)
cust_nonDefault <- c(cust <- c("Cust7","Cust8","Cust9","Cust10"))
points(x = CreditCardBalance_nonDefault, y = Age_nonDefault, col = "blue", pch = 2)
text(CreditCardBalance_nonDefault, Age_nonDefault, cust_nonDefault, pos = 1)

## finding the line with least deviation 

## 4.45063 * x + 1.94715 * y = 100 - 8.76217

intercept <- 100 / 1.94715
slope <- -4.45063 / 1.94715
abline(intercept, slope)

## Four samples have non-zero deviation. 
## The maximum deviation is 8.76217
```

## Question 3

(a) What will be the solution of the linear program if we let k1 = 2 and k2 = 1?

```{r q3a}
## q3a

## min 0 * X1 + 0 * X2 + 2 * a + 1 * b
## Subject to:
## (For default samples: )
## w1*10 + w2*24 + a ≥ 100 
## w1*25 + w2*35 + a ≥ 100 
## w1*3 + w2*40 + a ≥ 100
## w1*10 + w2*30 + a ≥ 100  
## w1*30 + w2*20 + a ≥ 100  
## w1*18 + w2*20 + a ≥ 100 
## (For non-default samples: )
## w1*2 + w2*28 - b ≤ 100 
## w1*5 + w2*21 - b ≤ 100
## w1*4 + w2*30 - b ≤ 100 
## w1*0 + w2*33 - b ≤ 100 
## a ≥ 0

cvec <- c(0, 0, 2, 1)
bvec <- c(100, 
          100, 
          100, 
          100, 
          100, 
          100, 
          -100, 
          -100, 
          -100, 
          -100)
names(cvec) <- c("X1", "X2", "a", "b")
names(bvec) <- c("Default1", 
                 "Default2", 
                 "Default3", 
                 "Default4", 
                 "Default5", 
                 "Default6", 
                 "Non-Default1", 
                 "Non-Default2", 
                 "Non-Default3", 
                 "Non-Default4")
Amat <- rbind(c(10, 24, 1, 0), 
              c(25, 35, 1, 0), 
              c(3, 40, 1, 0),
              c(10, 30, 1, 0),
              c(30, 20, 1, 0),
              c(18, 20, 1, 0),
              c(-2, -28, 0, 1),
              c(-5, -21, 0, 1),
              c(-4, -30, 0, 1),
              c(-10, -33, 0, 1))
operator <- c(">=",
              ">=",
              ">=",
              ">=",
              ">=",
              ">=",
              ">=",
              ">=",
              ">=",
              ">=")
res <- solveLP(cvec, bvec, Amat, const.dir = operator, maximum = FALSE)
print(res)
```

(b) How many samples are now misclassified?

```{r q3b}

CreditCardBalance <- c(10, 25, 3, 10, 30, 18)
Age <- c(24, 35, 40, 30, 20, 20)
cust <- c("Cust1","Cust2","Cust3","Cust4","Cust5","Cust6")

data <- data.frame(CreditCardBalance, Age)
plot(data, col = "red", pch = 1, xlim = c(0, 50), ylim = c(0, 50))
text(CreditCardBalance, Age, cust, pos = 1)

## Non default
CreditCardBalance_nonDefault <- c(2, 5, 4, 10)
Age_nonDefault <- c(28, 21, 30, 33)
cust_nonDefault <- c(cust <- c("Cust7","Cust8","Cust9","Cust10"))
points(x = CreditCardBalance_nonDefault, y = Age_nonDefault, col = "blue", pch = 2)
text(CreditCardBalance_nonDefault, Age_nonDefault, cust_nonDefault, pos = 1)

## finding the line with least deviation 

intercept <- 100 / 2.13415
slope <- -4.87805 / 2.13415
abline(intercept, slope)

## One sample has non-zero deviation. 
## The maximum deviation is 8.76217
```

(c) What would you do if you now think it is very important that all the non-default
customers must be correctly classified?

```{r q3c}
## Main goal -> Make sure deviation b becomes 0 

## Try out different values of constant k2 for b to achieve that 

## q3c

## min 0 * X1 + 0 * X2 + 1 * a + 2 * b
## Subject to:
## (For default samples: )
## w1*10 + w2*24 + a ≥ 100 
## w1*25 + w2*35 + a ≥ 100 
## w1*3 + w2*40 + a ≥ 100
## w1*10 + w2*30 + a ≥ 100  
## w1*30 + w2*20 + a ≥ 100  
## w1*18 + w2*20 + a ≥ 100 
## (For non-default samples: )
## w1*2 + w2*28 - b ≤ 100 
## w1*5 + w2*21 - b ≤ 100
## w1*4 + w2*30 - b ≤ 100 
## w1*0 + w2*33 - b ≤ 100 
## a ≥ 0

cvec <- c(0, 0, 1, 2)
bvec <- c(100, 
          100, 
          100, 
          100, 
          100, 
          100, 
          -100, 
          -100, 
          -100, 
          -100)
names(cvec) <- c("X1", "X2", "a", "b")
names(bvec) <- c("Default1", 
                 "Default2", 
                 "Default3", 
                 "Default4", 
                 "Default5", 
                 "Default6", 
                 "Non-Default1", 
                 "Non-Default2", 
                 "Non-Default3", 
                 "Non-Default4")
Amat <- rbind(c(10, 24, 1, 0), 
              c(25, 35, 1, 0), 
              c(3, 40, 1, 0),
              c(10, 30, 1, 0),
              c(30, 20, 1, 0),
              c(18, 20, 1, 0),
              c(-2, -28, 0, 1),
              c(-5, -21, 0, 1),
              c(-4, -30, 0, 1),
              c(-10, -33, 0, 1))
operator <- c(">=",
              ">=",
              ">=",
              ">=",
              ">=",
              ">=",
              ">=",
              ">=",
              ">=",
              ">=")
res <- solveLP(cvec, bvec, Amat, const.dir = operator, maximum = FALSE)
print(res)

## deviation for non default customers become 0, hence show that none of non default customers are incorrectly classified. 
```