---
title: "BT2103 A1 A0236437B"
author: "Lo Zhi Hao"
date: "2022-09-14"
output: html_document
---
For this assignment, you will experiment with the SVM classifier.
1. Download and install R

2. Install package e1071. 
```{r 2}
## install.packages("e1071")
library(dplyr)
library(e1071)
```

3. Download the Rice (Cammeo and Osmancik) Data Set (RiceDataset.txt) set from Luminus
Assignment1 folder. 

4. Load the data file RiceDataset.txt into R and call it rice.

```{r 4}
## download and load file 

rice <- read.table("RiceDataset.txt", header = TRUE)
head(rice)
summary(rice)
```

5. Retype the following lines of R commands:

```{r 5, echo = FALSE}
## copy down the r commands as instructed 

library(e1071)
index <- 1:nrow(rice)
set.seed(12345)
testindex <- sample(index, trunc(length(index)/3))
testset <- rice[testindex,]
trainset <- rice[-testindex,]


svm.model <- svm(as.factor(CLASS) ~ . , data = trainset,type="C-classification", kernel="linear")
svm.model
svm.model$SV
svm.model$index


results_train <- predict(svm.model,trainset[,-8])
results_test <- predict(svm.model, testset[,-8] )
table(pred=results_train,actual=trainset$CLASS)
table(pred=results_test,actual=testset$CLASS)
mean(results_train == trainset$CLASS)
mean(results_test == testset$CLASS)


svm.crossmodel <- svm(as.factor(CLASS) ~ . , data=rice, cross=10,type="C-classification",kernel="linear",cost=1)
results_rice <- predict(svm.crossmodel,rice[,-8])
table(pred=results_rice,actual=rice$CLASS)
mean(results_rice == rice$CLASS)


svm.weightedmodel <- svm(as.factor(CLASS) ~ . , data=rice, cross=10, type="C-classification",kernel="linear",cost=1, class.weights=c(Cammeo=0.9,Osmancik=0.1))
results_rice <- predict(svm.weightedmodel,rice[,-8])
table(pred=results_rice,actual=rice$CLASS)
mean(results_rice == rice$CLASS)

## weights improve after change(change to similar to real distribution)
## changing kernels do not help
## changing times of cross validation does not help
## Changing to type `nu-classfication` classification improves the model 
## Increasing cost (penalty on wrongly classified values) improves the model slightly

```

6. Deliverable: Upload a report as pdf file by the deadline into Luminus folder Assignment1-
Submission. Include in the report your answers to the following questions:


• What does each line of the R code above do? Explain any output that is produced by
the R command.

```{r line1 }

## I have created code chunks according to the main functionality each of the chunks are dedicated to doing 

## For the first chunk of R code, we are setting up the packages and classifying the existing data into training and test sets
library(e1071) # opening the e1071 package
index <- 1:nrow(rice) # Find the total number of rows and  create a vector with the number of rows in rice dataset
set.seed(12345) # set seed to make sure the pseudorandom number generator in following lines always generates the same output and helps in debugging
testindex <- sample(index, trunc(length(index)/3)) # takes a sample from the rice rows vectors that is 1/3 of the original size
testset <- rice[testindex,] # creating a test dataset to prepare for testing of the SVM model 
trainset <- rice[-testindex,] # creating a training dataset to prepare for the running of SVM model

## Second chunk of code is about running and printing out the results from the SVM model
svm.model <- svm(as.factor(CLASS) ~ . , data = trainset,type="C-classification", kernel="linear") # Running the SVM model using the svm function in R 
svm.model # Calling a summary of the svm model, 
svm.model$SV # printing out the resulting support vectors
svm.model$index # printing out the index of the resulting support vectors in the data matrix

## Utilising the svm model to predict and measure the accuracy of the model on test dataset
results_train <- predict(svm.model,trainset[,-8]) # generating prediction results from training data by using the svm model 
results_test <- predict(svm.model, testset[,-8]) # generating prediction results from test data by using the svm model
table(pred=results_train,actual=trainset$CLASS) # creating a contingency table to represent the confusion matrix of the model in the training dataset
table(pred=results_test,actual=testset$CLASS) # creating a contingency table to represent the confusion matrix of the model in the test dataset
mean(results_train == trainset$CLASS) # getting the accuracy of estimations of predicted results in training dataset (true positive, true negative)
mean(results_test == testset$CLASS) # getting the accuracy of estimations of predicted results in test dataset (true positive, true negative)

## Rerun the model using addtional cross validation and penalising parameters 
svm.crossmodel <- svm(as.factor(CLASS) ~ . , data=rice, cross=10,type="C-classification",kernel="linear",cost=1) # Rerun the svm model with more arguments (cross for k-times cross validation and cost for penalising constraint violations)
results_rice <- predict(svm.crossmodel,rice[,-8]) # generating prediction results from the entire rice dataset by using the svm model
table(pred=results_rice,actual=rice$CLASS) # creating a contingency table to represent the confusion matrix of the model in the test dataset
mean(results_rice == rice$CLASS) # getting the accuracy of estimations of predicted results in rice dataset (true positive, true negative)

## Rerun the model using additional weight parameter
svm.weightedmodel <- svm(as.factor(CLASS) ~ . , data=rice, cross=10, type="C-classification",kernel="linear",cost=1, class.weights=c(Cammeo=0.9,Osmancik=0.1)) # Rerun the svm model with more arguments (class.weight to take into account the assymetrical nature of cammeo and osmancik class size)
results_rice <- predict(svm.weightedmodel,rice[,-8]) # generating prediction results from the entire rice dataset by using the svm model
table(pred=results_rice,actual=rice$CLASS) # creating a contingency table to represent the confusion matrix of the model in the test dataset
mean(results_rice == rice$CLASS) # getting the accuracy of estimations of predicted results in rice dataset (true positive, true negative)

## cl <- rice %>% group_by(CLASS) %>% count()
## cl
```

• Provide some suggestions as how we can possibly obtain better results while using this
SVM package on the Rice data set.

Some possible changes to make to the SVM model in order to achieve better results includes:

1. increase the amount of penalty on incorrectly classified data by increasing the cost to better take into account incorrect data classification

2. using different kernels to take into account different kinds of possible mathematical relationships other than linear relationships

3. use different type of classification such as the `nu-classification` method 

4. Apply different weightage for the two rice classes (actual data is 1630 : 2180)